SunPy - Python for Solar Physics

### 2014-01-24 4-Util.tex
added code example for wcs package and tidying up of text

\section{Utilities}
SunPy is meant to provide a consistent environment for solar data analysis. In 
order to achieve this goal SunPy provides a number of utility functions which 
are used by the other SunPy modules and are made available to the user. Two 
modules are described in this Section, namely, solar coordinates (coords), and 
the Sun module (sun).

\subsection{Coordinates}
Coordinate transformations are frequently a necessary task within the solar 
data analysis workflow. Likely the most often used transformation is from 
observer coordinates (e.g. sky coordinates) to a coordinate system which is 
mapped onto the solar surface (e.g. latitude and longitude). This 
transformation is necessary to compare the true physical distance between 
different solar features. This type of transformation is not unique
to solar observations (e.g. planetary) but it is not provided by the Astropy 
coordinates package. The coordinates package implements the World Coordinate 
System (WCS) for solarcoordinates as described by \cite{Thompson2000}. The 
coordinate system transformations currently implemented as those most useful 
for most solar data analysis, namely converting from Helioprojective-Cartesian 
(HPC) to Heliographic coordinates (HG). HPC which describes the positions on 
the Sun as angles measured from the center of the solar disk (usually in 
arcseconds) using cartesian coordinates (X, Y) This is the coordinates 
most often defined in solar image data (e.g. SDO/AIA images, SOHO/EIT, TRACE). 
HG coordinates express positions on the Sun using longitude and latitude on 
the 
solar sphere. There are two standards for this coordinate system, 
Stonyhurst-Heliographic where the origin is at the intersection of the solar 
equator and the central meridian as seen from Earth and 
Carrington-Heliographic 
is a system which is fixed to the Sun and does not depend on Earth. The 
implementation of these transformations pass through common coordinate system 
called Heliocentric-Cartesian (HCC) where positions are expressed in true 
(deprojected) physical distances instead of angles on the celestial sphere.
These transformations require some knowledge of the location of the observer 
which is usually provided by the image header information. In case where it is 
not, the observer is assumed to be at Earth.

\subsubsection{Sun}
The purpose of this module is to provide solar-specific data

\subsubsection{constants}
The purpose of the constants package is to provide a number of solar-related 
constants in order to provide consistency in the calculations of derived solar 
values both within the SunPy code base but also to the user. Every solar 
constants is provided as a Constant objects as defined by Astropy. Each 
Constant object defines a Quantity, a number associated with a unit, along with 
the constant's provenance (i.e. reference) as well as its uncertainty. Using 
Astropy's Quantities, any solar constant can easily be converted between 
different units as can be seen in Figure~\ref{fig:constants_code} as well
as between SI or cgs with a single command.  As these objects inherit from 
Numpy's ndarray, they work well with standard representations of numbers.
For convenience a number shortcuts to frequently used constants are provided 
directly when importing the sub-package. A larger list of constants can be 
accessed through an interface modelled on that provided by the SciPy constants 
package and are available as a dictionary called \texttt{physical\_constants}. 
To view all quickly a \texttt{print\_all()} function is available.

\begin{listing}[h]
\begin{minted}{pycon}

@@

\end{minted}
\caption{Using sunpy.sun.constants sub-package.}
\label{code:constants_code}
\end{listing}

>>>>

\section{Utilities}
SunPy is meant to provide a consistent environment for solar data analysis. In 
order to achieve this goal SunPy provides a number of utility functions which 
are used by the other SunPy modules and are made available to the user. Two 
modules are described in this Section, namely, solar coordinates (wcs), and 
the Sun module (sun).

\subsection{WCS Coordinates}
Coordinate transformations are frequently a necessary task within the solar 
data analysis workflow. Likely the most often used transformation is from 
observer coordinates (e.g. sky coordinates) to a coordinate system which is 
mapped onto the solar surface (e.g. latitude and longitude). This 
transformation is necessary to compare the true physical distance between 
different solar features. This type of transformation is not unique
to solar observations (e.g. planetary) but is not often considered by astronomical packages
such as the Astropy 
coordinates package. The wcs package implements the World Coordinate 
System (WCS) for solar coordinates as described by \cite{Thompson2000}. The 
coordinate system transformations currently implemented are those most useful 
for most solar data analysis, namely converting from Helioprojective-Cartesian 
(HPC) to Heliographic coordinates (HG). HPC describes the positions on 
the Sun as angles measured from the center of the solar disk (usually in 
arcseconds) using cartesian coordinates (X, Y). This is the coordinates 
most often defined in solar image data (e.g. SDO/AIA images, SOHO/EIT, TRACE). 
HG coordinates express positions on the Sun using longitude and latitude on 
the solar sphere. There are two standards for this coordinate system, 
Stonyhurst-Heliographic where the origin is at the intersection of the solar 
equator and the central meridian as seen from Earth and 
Carrington-Heliographic is a system which is fixed to the Sun and does not depend on Earth. The 
implementation of these transformations pass through a common coordinate system 
called Heliocentric-Cartesian (HCC) where positions are expressed in true 
(deprojected) physical distances instead of angles on the celestial sphere.
These transformations require some knowledge of the location of the observer 
which is usually provided by the image header . In case where it is 
not, the observer is assumed to be at Earth. An example of the

\begin{listing}[H]
\begin{minted}{pycon}
>>> from sunpy import wcs
Find the helioprojective coordinate of a given longitude and latitude on the Sun
>>> wcs.convert_hg_hpc(10,53)
(100.49244115330731, 767.97438321917502)
Convert that position back to heliographic coordinates
>>> wcs.convert_hpc_hg(100.49, 767.97)
(9.9996521808465175, 52.999563684874893)
Try to convert a position which is not on the Sun to heliographic coordinates
>>> wcs.convert_hpc_hg(-1500,0)
sunpy/wcs/wcs.py:180: RuntimeWarning: invalid value encountered in sqrt
  distance = q - np.sqrt(distance)
(nan, nan)
Convert sky coordinate to a position on the Sun in cartesian coordinates (HCC)
>>> wcs.convert_hpc_hcc(-300,400, z = True)
(-216716967.63331246, 288956420.9477042, 594364636.2208252)
\end{minted}
\caption{Using the sunpy.wcs sub-package.}
\label{code:wcs_code}
\end{listing}

\subsubsection{Sun}
The purpose of this module is to provide solar-specific data for example ephemeris and
solar constants. The main namespace contains a number of functions which provide solar
ephemeris such as the sun to earth distance, solar cycle number, the mean anomaly, etc.
All of these functions take a time as their input which can be provided in any 
parse\_time
compatible format. 
A future version of SunPy will integrate with the AstroPy coordinates package and will 
likely make use of an ephemeris package.

\subsubsection{constants}
The purpose of the constants package is to provide a number of solar-related 
constants in order to provide consistency in the calculations of derived solar 
values both within the SunPy code base but also to the user. Every solar 
constants is provided as a Constant objects as defined by Astropy. Each 
Constant object defines a Quantity, a number associated with a unit, along with 
the constant's provenance (i.e. reference) as well as its uncertainty. Using 
Astropy's Quantities, any solar constant can easily be converted between 
different units as can be seen in Figure~\ref{fig:constants_code} as well
as between SI or cgs with a single command.  As these objects inherit from 
Numpy's ndarray, they work well with standard representations of numbers.
For convenience a number of shortcuts to frequently used constants are provided 
directly when importing the sub-package. A larger list of constants can be 
accessed through an interface modelled on that provided by the SciPy constants 
package and are available as a dictionary called \texttt{physical\_constants}. 
To view them all quickly a \texttt{print\_all()} function is available.

\begin{listing}[H]
\begin{minted}{pycon}

@@

\end{minted}
\caption{Using the sunpy.sun.constants sub-package.}
\label{code:constants_code}
\end{listing}
